### 2015-10-30 postdoctoral/gamma_functions.tex
WIP compact lex prod proof
nailed the answer for \gamma

  Note that \(\check L\)
  is not a ``compactification'' as \(L\) does not necessarily
  embed as a dense subspace of \(\check L\): if \(L=I\), we might attempt to embed
  \(t\mapsto [0,t)\), but then note that the subspace topology induces the
  reverse Sorgenfrey interval as \(([0,s),[0,t])=([0,s),[0,t)]\) is open.
  However \(\hat L\) is
  the typical way of compactifying a linearly ordered space \(L\),
  provided \(L\) lacks a least element (otherwise the empty set is an [easily
  removable] isolated point in \(\hat L\)).

@@

  \begin{proof}
    Since \(\check\alpha=\alpha+1\) (actually equals, not just homeomorphic!),
    we get \(\varprojlim^\star \{2, \gamma, \alpha\}
      \cong \check\alpha =
    \alpha+1\) for free.
  \end{proof}

@@

  \begin{proof}
    If \(\alpha\) is finite, then of course all (leftward) sets are
    closed and we get \(\hat\alpha=\check\alpha=\alpha+1\) for free.
    Otherwise, since \(\alpha\) lacks a greatest point, \(\hat\alpha\)
    is homeomorphic to its usual compactification \(\alpha+1\).
  \end{proof}

@@

  The ready may verify the following examples:

  \begin{example}
    \(\varprojlim\{2,\gamma,I\}\cong \check L\lexTimes 1\)
  \end{example}
  \begin{example}
    \(\varprojlim\{I,\gamma,I\}\cong I\lexTimes I\)
  \end{example}
  \begin{example}
    \(\varprojlim\{I,\gamma,\omega\}\cong (\omega\lexTimes[0,1))\cup\{\infty\}\)
  \end{example}
  \begin{example}
    \(\varprojlim\{I,\gamma,\omega+1\}\cong \omega+1\lexTimes I\)
  \end{example}
  \begin{example}
    \(\varprojlim\{I,\gamma,-\omega\}\cong \{-\infty\}\cup(-\omega\lexTimes(0,1])\)
  \end{example}
  \begin{example}
    \(\varprojlim\{I,\gamma,I\}\cong I\lexTimes I\)
  \end{example}

  \begin{theorem}
    If \(K,L\) are compact linearly orders, then
    \(\varprojlim\{K,\gamma,L\}\cong L\lexTimes K\).
  \end{theorem}

  \begin{proof}

  \end{proof}

>>>>

  Note that \(\check L\)
  is not a ``compactification'' as \(L\) does not necessarily
  embed as a dense subspace of \(\check L\): if \(L=I\), we might attempt to embed
  \(t\mapsto [0,t)\), but then note that the subspace topology induces the
  reverse Sorgenfrey interval as \(([0,s),[0,t])=([0,s),[0,t)]\) is open.
  However \(\hat L\) is
  the typical way of compactifying a linearly ordered space \(L\),
  provided \(L\) lacks a least element (otherwise the empty set is an [easily
  removable] isolated point in \(\hat L\)). Note that we \textbf{always}
  assume that \(\emptyset\in\hat L\):

  \begin{example}
    \(\hat I\cong\{-\infty\}\cup I\) where
    \(\emptyset\mapsto-\infty\) and \([0,t]\mapsto t\).
  \end{example}

  \begin{example}
    For limit ordinals \(\alpha\), \(\hat \alpha\cong \alpha+1\),
    and for all other infinite ordinals, \(\hat\alpha\cong\alpha\).
    (The addition of a new least isolated point is of course irrelevant).
  \end{example}

@@

  \begin{proof}
    Since \(\check\alpha=\alpha+1\) (actually equals, not just homeomorphic!),
    we get \(\varprojlim^\star \{2, \gamma, \alpha\}
      \cong \check\alpha =
    \alpha+1\) for free.
    Note that C and Varagona used this in (TODO create citataion) to break
    metrizability in uncountable-ordinal-indexed inverse limits (for any V-map
    there exists a two-point set \(2\) such that \(f\rest 2\supseteq\gamma\),
    that is, ``\(f\) has condition \(\Gamma\)'').
  \end{proof}

  We may generalize theorem 8 as follows:

  \begin{theorem}
    If \(M\) is a LOTS with minimum \(0\) and maximum \(1\),
    then \(\varprojlim\{M,\gamma,L\}\cong \hat L\lexTimes M/\sim\),
    where \(\<(\leftarrow,l_0],1\>\sim\<(\leftarrow,l_1],0\>\) if
    \(l_0<l_1\) and \((l_0,l_1)=\emptyset\),
    and where \(\<A,m\>\sim\<A,m'\>\) if \(A\in\hat L\setminus L\).
  \end{theorem}

  \begin{proof}
    Let \(\rho(\vect x)=\cl\{l\in L:\vect x(l)>0\}\), \(v(\vect 0)=0\),
    and \(v(\vect x)=\min\{\vect x(l):l\in\rho(\vect x)\}\) otherwise.
    Say \(\vect x<\vect y\) if \(\rho(\vect x)\subsetneq\rho(\vect y)\)
    or both \(\rho(\vect x)=\rho(\vect y)\) and \(v(\vect x)<v(\vect y)\).
    The reader may verify that this is a linear order on
    \(\varprojlim\{M,\gamma,L\}\), and
    \(\theta(\vect x)=\<\rho(\vect x),v(\vect x)\>\in\hat L\lexTimes M/\sim\)
    preserves order.
    For each left-closed set \(A\) and \(m\in M\),
    let \(\vect x_{A,m}(l)=1\) for \(l\in A\) unless \(l\) is the supremum
    element of \(A\),
    \(\vect x_{A,m}(l)=m\) if \(l\) is the supremum of \(A\),
    and \(\vect x_{A,m}(l)=0\) for \(l\not\in A\).
    To complete the proof, we should demonstrate that the linear order we
    defined induces the topology of the inverse limit, and that \(\theta\)
    is a surjection.

    A basic open set in \(\varprojlim\{M,\gamma,L\}\subseteq L^m\)
    is of the form
    \([U,F]\) where \(U(l)\) is an open interval in \(M\) for each
    \(l\in F\in [L]^{<\omega}\), and
    \([U,F]=\{\vect x:l\in F\Rightarrow\vect x(l)\in U(l)\}\).
    If we assume that \([U,F]\) is non-empty, one of the following
    must hold:

    \begin{itemize}
      \item
        \(U[l_0]=(a,b)\) for some \(l_0\in F\).
        Then \([U,F]=[U,\{l_0\}]\), and note that
        \(
          [U,\{l_0\}]
            =
          (\vect x_{(\leftarrow,l_0],a},\vect x_{(\leftarrow,l_0],b})
        \).
      \item
        \(U(l_0)=(a,1]\) and \(U(l_1)=[0,b)\) for some \(l_0<l_1\in L\)
        and \([U,F]=[U,\{l_0,l_1\}]\). Then
        \([U,\{l_0,l_1\}]=(\vect x_{l_0,a},\vect x_{l_1,b})\).
    \end{itemize}
    In the other direction, consider \(\vect y\in(\vect x,\vect z)\).
    \begin{itemize}
      \item
        In the case that \(l_0\in\rho(\vect y)\setminus\rho(\vect x)\)
        and \(l_1\in\rho(\vect z)\setminus\rho(\vect y)\), let
        \(U(l_0)=(0,1]\), \(U(l_1)=[0,v(\vect z))\) and note
        \(\vect y\in[U,\{l_0,l_1\}]\subseteq(\vect x,\vect z)\).
      \item
        In the case that \(l_0\in\rho(\vect y)\setminus\rho(\vect x)\),
        \(\rho(\vect y)=\rho(\vect z)\), and \(v(\vect y)<v(\vect z)\),
        it follows that \(\rho(\vect y)=\rho(\vect z)=(\leftarrow,l_1]\),
        so let \(U(l_0)=(0,1]\), \(U(l_1)=[0,v(\vect z))\) and note
        \(\vect y\in[U,\{l_0,l_1\}]\subseteq(\vect x,\vect z)\).
      \item
        In the case that
        \(\rho(\vect x)=\rho(\vect y)\), \(v(\vect x)<v(\vect y)\),
        and \(l_1\in\rho(\vect z)\setminus\rho(\vect y)\),
        it follows that \(\rho(\vect x)=\rho(\vect y)=(\leftarrow,l_0]\),
        so let \(U(l_0)=(v(\vect x),1]\), \(U(l_1)=[0,v(\vect z))\) and note
        \(\vect y\in[U,\{l_0,l_1\}]\subseteq(\vect x,\vect z)\).
      \item
        In the case that
        \(\rho(\vect x)=\rho(\vect y)=\rho(\vect z)\) and
        \(v(\vect x)<v(\vect y)<v(\vect z)\),
        it follows that
        \(\rho(\vect x)=\rho(\vect y)=\rho(\vect z)=(\leftarrow,l_0]\),
        so let \(U(l_0)=(v(\vect x),v(\vect z))\) and note
        \(\vect y\in[U,\{l_0\}]=(\vect x,\vect z)\).
    \end{itemize}

    We conclude by
    showing that \(\theta\) is a surjection. If \(B\in\hat L\setminus L\)
    and \(m\in M\), consider \(\<B,m\>\).
    \(B\) lacks a supremum in \(L\), so
    \(\vect x_{B,0}(l)=1\) for \(l\in B\) and \(\vect x_{B,0}(l)=0\) otherwise.
    So \(\theta(\vect x_{B,0})=\<\cl B,1\>=\<B,1\>\sim\<B,m\>\) for all
    \(m\in M\). Otherwise, \(B=(\leftarrow,l_1]\) for some \(l_1\in L\).
    Let \(m>0\).
    Then
    \(
      \theta(\vect x_{(\leftarrow,l_1],m})
        =
      \<\cl (\leftarrow,l_1],v(\vect x_{(\leftarrow,l_1],m})\>
        =
      \<(\leftarrow,l_1],m\>
    \).
    Finally, we want to map onto \(\<(\leftarrow,l_1],0\>\).
    If there exists \(l_0<l_1\) with \((l_0,l_1)=\emptyset\), then
    \(
      \theta(\vect x_{(\leftarrow,l_1],0})
        =
      \theta(\vect x_{(\leftarrow,l_0],1})
        =
      \<(\leftarrow,l_0],1\>
    \)
    will suffice. Otherwise,
    \(
      \theta(\vect x_{(\leftarrow,l_1],0})
        =
      \<\cl (\leftarrow,l_1),v(\vect x_{(\leftarrow,l_1),0})\>
        =
      \<(\leftarrow,l_1],0\>
    \).
  \end{proof}

  Here are some applications:

  \begin{example}
    \(
      \varprojlim\{2,\gamma,I\}
        \cong
      (\hat I\setminus\emptyset)\lexTimes 2
        \cong
      I\lexTimes 2
        \cong
      \check I
    \) (of course, this could be found quicker with theorem 8).
  \end{example}

  \begin{example}
    \(
      \varprojlim\{I,\gamma,I\}
        \cong
      (\hat I\setminus\emptyset)\lexTimes I
        \cong
      I\lexTimes I
    \).
  \end{example}

  \begin{example}
    For infinite ordinals \(\alpha\),
    \(
      \varprojlim\{I,\gamma,\alpha\}
        \cong
      (\alpha\lexTimes[0,1))\cup\{\infty\}
    \).
    In particular, \(\alpha=\kappa\) for an infinite cardinal \(\kappa\)
    gives the closed long ray of length \(\kappa\).
  \end{example}

@@

  \begin{proof}
    If \(\alpha\) is finite, then of course all (leftward) sets are
    closed and we get \(\hat\alpha=\check\alpha=\alpha+1\) for free.
    Otherwise, as observed previously \(\hat\alpha\)
    is homeomorphic to its usual compactification \(\alpha+1\) for
    limit ordinals.
  \end{proof}

@@
